---
title: "For Botanist"
author: "Stephen Brownsey"
date: "03/11/2019"
output: pdf_document
---

## Introduction
The code in this .rmd was used for the discussion with the botanist on the 03/11/2019. It also contains extra EDA on top of the EDA already undertaken as part of the exploratory_data.rmd file
```{r libraries, results= "hide", message=F, warning=F, echo = FALSE}
library(tidyverse)
library(ggmap)
library(leaps)
library(GGally)
library(gridExtra)
library(factoextra)
library(FactoMineR)
library(BBmisc)
library(pvclust)
library(NbClust)
library(class)
library(cluster)
library(clValid)
library(modeest)
```

### Map Analysis
This is an image of where each of the orchards is located, more for interest but would you think that they would go around taking readings of different orchards based on similar days when they are close to each other? If so, would it make sense to apply some models to just one area?
```{r, echo = FALSE, warning=FALSE}
#Loading the data in from paper ~ downloaded from their resources as a .xlsx, created csv with useful page for quicker reading.
data <- read.csv("data.csv") %>%
  #renaming as column name had a few unrecognised characters in front
  rename(orchard = contains("orchard"))

#Creating an indicator variable to show whether the Orchard appeared in both years or just 1 (0 -> No, 1 -> Yes)  
years <- data %>% 
  select(orchard, year) %>%
  distinct() %>%
  count(orchard) %>%
  rename(both_years = n) %>%
  mutate(both_years = (both_years - 1)) %>%
  mutate(id = row_number()) 


data <- merge(data, years)%>%
  na.omit()

#Isolating the lat and long
lat_long <- data %>% 
  select(lat,long,both_years) %>%
  distinct()

register_google(key = "AIzaSyB9Hpt0vTWrALpm0iUyJxW6C2IuHZylpC8")
map <- get_map(location = c(lon = mean(lat_long$lon), lat = mean(lat_long$lat)), zoom = 9,
                      maptype = "satellite", scale = 2)

#~~~~~Plotting the locations of all the orchards:
plotted_map <- ggmap(map) +
  geom_point(data = lat_long, aes(x = long, y = lat, fill = both_years , alpha = 0.8), size = 5, shape = 21) +
  guides(fill = FALSE, alpha = FALSE, size = FALSE) +
  ggtitle("Orchard Locations") +
  xlab("Longitude") + 
  ylab("Latitude")
plotted_map
## 
```

The variables in general are very correlated, as you would expect, but this means that for the decision models I would want to be looking at a subset of these but from a botanist's point of view which variables would she consider best?
Not that clear in a .pdf but lots are upwards of 0.8/0.9 correlated.

### GGTally
Takes about 5-10 mins to run - produces all the various correlation and associated plots between variables
```{r, echo = FALSE}
#bees_data <- ggpairs(data = data, columns = 7:13, title = "Bees Data")
#bees_data
#predictor_data <- ggpairs(data = data, columns = 13:34, title = "Predictor Data")
#predictor_data
```

### Looking at individual variables
```{r, echo = FALSE, warning=FALSE}
day_data <- data %>%
  filter(day %in% c("1","2")) %>%
  mutate(year_day = ifelse(str_detect(day, "1") & str_detect(year, "3"), '11', 
                           ifelse(str_detect(day, "2") & str_detect(year, "3"), '12',
                           ifelse(str_detect(day, "1") & str_detect(year, "4"), '21',
                            ifelse(str_detect(day, "2") & str_detect(year, "4"), '22', "55"))))) %>%
  mutate(year_day = factor(year_day, labels = c("Year 3 and Day 1","Year 3 and Day 2", "Year 4 and Day 1", "Year 4 and Day 2"))) %>%
  na.omit()


```

### Code post 25/10
Violin plots of the data - I am a fan of what these show, but in particular note the one for bloom.index (at the bottom), I had thought that day 1 would be < 50% bloom and day 2 post 50% bloom but this clearly shows this is not the case - thoughts? Have also included a distributions plot of the variables in the dataset for interest as it could be helpful when thinking about these things.
```{r, echo = FALSE, warning=FALSE}
data_2012 <- day_data %>%
  filter(year == 4) 

#Looking at violin plots of the data
violin_data <- day_data %>%
  select(c("region":"X2000nat")) %>%
  gather(key = "explanatory", value = "value",-region, -day) %>%
  na.omit()

violin_plot <- function(x, xlab, ylab, title){
 x %>% ggplot() +
    geom_violin(aes(x = factor(explanatory), y = value, fill = explanatory, colour = explanatory)) +
    ylab(ylab) + 
    xlab(xlab) +
    ggtitle(title) +
    theme(legend.position = "none") +
    theme_classic()
}
 
violin_plot_bees <- violin_data %>%
  subset(explanatory %in% colnames(day_data[4:12])) %>%
  violin_plot("Bee Factor", "Bee Rating", "Distributions of each Bee Variable")
violin_plot_bees


violin_plot_fungicides <- violin_data %>%
  subset(explanatory %in% colnames(day_data[c(19,22,23,24)])) %>%
  violin_plot("Fungicide Factor","Fungicide Rating","Distributions of each Fungicide Variable")
violin_plot_fungicides

violin_plot_insecticide <- violin_data %>%
  subset(explanatory %in% colnames(day_data[c(20,21,25:28)])) %>%
  violin_plot("Insecticide Factor", "Insecticide Rating", "Distributions of each Insecticide Variable")

#Distributions of all variables
distributions <- violin_data %>%
  ggplot() +
  geom_density(aes(value)) +
  facet_wrap(~explanatory, scales = "free") +
  theme_classic()
distributions

#Bloom values - particularly not ideal as it looks like the day 1 / day 2 actually is just random
#lol not even in relation to bloom levels 
bloom_plot <- day_data %>%
  group_by(region, day) %>%
  mutate(group = paste(region, day)) %>%
  ggplot(aes(x= group, y = bloom.index)) +
  geom_violin(aes(fill = group, colour = group)) +
  geom_jitter(height = 0, width = 0.05) + 
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "Region and Day", y = "Bloom Index", title = "Violin Plot of Bloom Index by Region and Day")
bloom_plot


bloom_plot_2012 <- data_2012 %>%
  group_by(region, day) %>%
  mutate(group = paste(region, day)) %>%
  ggplot(aes(x= group, y = bloom.index)) +
  geom_violin(aes(fill = group, colour = group)) +
  geom_jitter(height = 0, width = 0.05) + 
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "Region and Day", y = "Bloom Index", title = "Violin Plot of Bloom Index by Region and Day")
bloom_plot_2012
```


## Graphics from botanists chat
Extra graphics and analysis undertaken based on the chat with the botanist to see relationships.
```{r, echo = FALSE, warning=FALSE}
group_day_data <-  day_data %>%
  group_by(region, day) %>%
  mutate(group = paste(region, day))


bloom_plot <- group_day_data  %>%
  ggplot(aes(x= orchard, y = bloom.index, colour = group, shape = group)) +
  geom_point()+
  theme_classic() +
  labs(x = "Region and Day", y = "Bloom Index", title = "Scatter_plot of Bloom Index by Region and Day")
bloom_plot


bloom_plot_2012 <- group_day_data  %>%
  ggplot(aes(x= orchard, y = bloom.index, colour = group, shape = group)) +
  geom_point()+
  theme_classic() +
  labs(x = "Region and Day", y = "Bloom Index", title = "Scatter_plot of Bloom Index by Region and Day for 2012 Data")
bloom_plot_2012

by_day <- tibble(day_1 = data_2012 %>% 
                      filter(day == 1) %>%
                      select(bloom.index), 
                    day_2 = data_2012 %>% 
                      filter(day == 2) %>%
                      select(bloom.index), 
                    colour = data_2012 %>%
                      filter(row_number() %% 2 == 1) %>%
                      select(region)) 


day_1_2 <- by_day %>%
  ggplot() + 
  geom_point(aes(x = day_1$bloom.index, y = day_2$bloom.index, colour = colour$region)) +
  theme_bw()
day_1_2
```


The first step for us is to understand the chemicals applications better, including the combinations. The variables in the data are  scores developed by the authors that capture the impact the have, so that is quite ready to use and comparable, need to further investigate this.
```{r, echo = FALSE, warning=FALSE}
fungicide_vs_insecticide <- function(data,y,y1,  title){
  data %>%
    ggplot(aes(x = orchard)) +
  geom_point(aes_string( y = y), colour = "red") +
  geom_point(aes_string( y = y1), colour = "blue") + 
  theme_bw() +
  ggtitle(title) +
  ylab("Value")
}

data_2012 %>%
  fungicide_vs_insecticide("eiqB11F.pre","eiqB11I.pre","eiqB11F.pre in red, eiqB11I.pre in blue")

data_2012 %>%
  fungicide_vs_insecticide("eiqB11F.blm","eiqB11I.blm","eiqB11F.blm in red, eiqB11I.blm in blue")

data_2012 %>%
  fungicide_vs_insecticide("eiqB11F.pos","eiqB11I.pos","eiqB11F.pos in red, eiqB11I.pos in blue")+
  geom_point(aes(x = orchard, y = eiqB11I.pos.np), colour = "purple") 

density_plot <- function(y){
  data_2012 %>%
     ggplot() +
  geom_density(aes_string(y)) +
    theme_bw() +
    ggtitle(paste0("Density of ", y))
}


hist_plot <- function(y){
  data_2012 %>%
    ggplot(aes( fill = region)) +
    geom_histogram(aes_string(y)) + 
    theme_bw() +
    ggtitle(paste0("Histogram of " , y))
}

l <- c("eiqB11F.pos","eiqB11I.pos","eiqB11F.pre", "eiqB11I.pre","eiqB11F.blm", "eiqB11I.blm")
lapply(l, density_plot)
lapply(l, hist_plot)
#couple of examples for progress report
ex1 <- density_plot("eiqB11I.pos")
ex2 <- hist_plot("eiqB11I.blm")

#a <- grid.arrange(temperature_plot, honey_temp, ex1,ex2, nrow = 2)
``` 
Looking at kmeans for clustering of orchards to decide on decision model, using Kmeans for now to generate the outputs quickly as an EDA approach to clustering.

```{r, echo = FALSE, warning=FALSE}
#Can probably put all kmeans into this datatype and sort it that way
set.seed(666)
kmeans_clusters <- tibble(cluster_2_pre = kmeans(data_2012 
                                    %>% select(eiqB11F.pre, eiqB11I.pre), 2)$cluster,
                          cluster_3_pre = kmeans(data_2012 
                                    %>% select(eiqB11F.pre, eiqB11I.pre), 3)$cluster,
                          cluster_4_pre = kmeans(data_2012 
                                    %>% select(eiqB11F.pre, eiqB11I.pre), 4)$cluster,
                          cluster_2_blm = kmeans(data_2012 
                                    %>% select(eiqB11F.blm, eiqB11I.blm), 2)$cluster,
                          cluster_3_blm = kmeans(data_2012 
                                    %>% select(eiqB11F.blm, eiqB11I.blm), 3)$cluster,
                          cluster_4_blm = kmeans(data_2012 
                                    %>% select(eiqB11F.blm, eiqB11I.blm), 4)$cluster,
                          cluster_2_pos = kmeans(data_2012 
                                    %>% select(eiqB11F.pos, eiqB11I.pos), 2)$cluster,
                          cluster_3_pos = kmeans(data_2012 
                                    %>% select(eiqB11F.pos, eiqB11I.pos), 3)$cluster,
                          cluster_4_pos = kmeans(data_2012 
                                    %>% select(eiqB11F.pos, eiqB11I.pos), 4)$cluster,
                 )
# With this method only 7 out of the possible 16 arms in the decision model with have any results
current <- kmeans_clusters %>%
  select(cluster_4_pre, cluster_2_blm, cluster_2_pos) %>%
  group_by_all() %>% 
  summarise(COUNT = n())
current

current_2 <- kmeans_clusters %>%
  select(cluster_2_pre, cluster_2_blm, cluster_2_pos) %>%
  group_by_all() %>% 
  summarise(COUNT = n())
current_2

current_3 <- kmeans_clusters %>%
  select(cluster_3_pre, cluster_3_blm, cluster_3_pos) %>%
  group_by_all() %>% 
  summarise(COUNT = n())
current_3



#####Looking deeper into the clustering of the 2_2_2
left <-  kmeans_clusters %>%
  group_by_all() %>%
  mutate(id = row_number())


right <- data_2012 %>%
  group_by(orchard,day)  %>%
  mutate(id = row_number())
orchard_clusters <- bind_cols(left, right)

#Function to calulate the diff percentages of clusters
bloomarise <- function(x){
  x %>%
      summarise(fung_pre = round(mean(eiqB11F.pre), 2),
            insect_pre = round(mean(eiqB11I.pre), 2),
            fung_blm = round(mean(eiqB11F.blm), 2),
            insect_blm = round(mean(eiqB11I.blm), 2),
            fung_pos = round(mean(eiqB11F.pos), 2),
            insect_pos = round(mean(eiqB11I.pos), 2))
}


bloom_2_2_2 <- orchard_clusters %>%
  group_by(cluster_2_pre, cluster_2_blm, cluster_2_pos) %>%
  bloomarise
bloom_2_2_2

bloom_4_2_2 <- orchard_clusters %>%
  group_by(cluster_4_pre, cluster_2_blm, cluster_2_pos) %>%
  bloomarise
bloom_4_2_2

bloom_3_3_3 <- orchard_clusters %>%
  group_by(cluster_3_pre, cluster_3_blm, cluster_3_pos) %>%
  bloomarise
bloom_3_3_3




data_clusters <- tibble(kmeans_clusters, data_2012)
data_clusters %>%
  ggplot() +
  geom_point(aes(x = data_2012$orchard, y = data_2012$eiqB11I.pre,colour = factor(kmeans_clusters$cluster_4_pre))) + 
  geom_point(aes(x = data_2012$orchard, y = data_2012$eiqB11F.pre,colour = factor(kmeans_clusters$cluster_4_pre))) +
  theme_bw()

k_clusters <- tibble(by_day, kmeans = data_clusters %>% filter(row_number() %% 2 == 1))
pre <- k_clusters %>%
  ggplot() + 
  geom_point(aes(by_day$day_1$bloom.index, y = by_day$day_2$bloom.index, colour = factor(kmeans$kmeans_clusters$cluster_4_pre))) +
  labs(title = "Pre-Bloom Decisions", x = "Day 1 Bloom", y = "Day 2 Bloom" , colour='Decision Arm') + 
  theme_bw()
pre

blm <- k_clusters %>%
  ggplot() + 
  geom_point(aes(by_day$day_1$bloom.index, y = by_day$day_2$bloom.index, colour = factor(kmeans$kmeans_clusters$cluster_2_blm))) +
  labs(title = "During-Bloom Decisions", x = "Day 1 Bloom", y = "Day 2 Bloom" , colour='Decision Arm') + 
  theme_bw()
blm

pos <- k_clusters %>%
  ggplot() + 
  geom_point(aes(by_day$day_1$bloom.index, y = by_day$day_2$bloom.index, colour = factor(kmeans$kmeans_clusters$cluster_2_pos))) +
  labs(title = "Post-Bloom Decisions", x = "Day 1 Bloom", y = "Day 2 Bloom" , colour='Decision Arm') + 
  theme_bw()
pos
```

Looking at correlations of the bee variables
```{r}
a <- data.frame(matrix(0, nrow = 0, ncol = 0))

for(i in 1:7){
  for(j in 1:7){
  a[i,j] <- round(cor(data_2012[5+i], data_2012[5 + j]), 2)
  }
}
a
for(i in 1:7){
 names(a)[i] <- colnames(data_2012[5 +i])
}

```
### Objective 1b/c/1d further models and analysis
After speaking with botanist, no predefined appraoch will be taken.
All the models will be created using a data driven approach, so going to look at some different clustering methods fitted using the 19 orchards as the data and then compare the different models and clustering methods.
From the Kmeans analysis above, the different seed returning differing clusterings is not ideal, so some different approaches will be looked at. It is also clear that more factors need to be taken into account in the clustering process rather than just the insecticide and pesticide levels.

#### Version 2

Taking the above method a little further to look at different distance metric and add in X2000nat as another variable.

```{r}
cluster_2_data <- data_2012 %>%
  select(ends_with(".pre"),ends_with(".blm"), ends_with(".pos"), X2000nat)
#Shows no difference here although from previous work ward method was best
# matrix of methods to compare
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")
distances <- c("euclidean", "maximum", "manhattan", "canberra", 
               "binary", "minkowski")
names(distances) <- c("euclidean", "maximum", "manhattan", 
                      "canberra", "binary", "minkowski")
clust_comps <- matrix(nrow = length(distances), ncol = length(m), 
                      dimnames = list(distances,m))
# function to compute coefficient
ac <- function(distance, linkage) {
  dista <- dist(cluster_2_data , method = distance)
  #Agglomerative Nesting form of Hierarchical Clustering
  agnes(dista, method = linkage)$ac
}
for(i in 1:length(distances)) {
  for(j in 1:length(m)) {
    clust_comps[i,j] <- ac(distances[i], m[j])
  }
}


#Clusters of whole data combined - 7
fviz_nbclust(NbClust(cluster_2_data, distance="euclidean", 
                                min.nc=2, max.nc=4, method="ward.D2", index="all"))
#pre ~ max of 4 clusters at each time point in method, but  in general if you allow
#More as this argument then more returned a better  optimal cluster number
fviz_nbclust(NbClust(cluster_2_data %>% select(ends_with(".pre"), X2000nat), distance="euclidean", 
                                min.nc=2, max.nc=4, method="ward.D2", index="all"))
#during
fviz_nbclust(NbClust(cluster_2_data %>% select(ends_with(".blm"), X2000nat), distance="euclidean", 
                                min.nc=2, max.nc=4, method="ward.D2", index="all"))
#post
fviz_nbclust(NbClust(cluster_2_data %>% select(ends_with(".pos"), X2000nat), distance="euclidean", 
                                min.nc=2, max.nc=4, method="ward.D2", index="all"))
```


Step 1: shows for all the data and pre bloom stage = 3 clusters is best out of our cluster range considered. 
```{r}
#Applying agglomerative clustering with 3 clusters:
clustered <- agnes(dist(cluster_2_data, method = "euclidian"), 
                   diss=TRUE, method = "ward")
# add cluster labels to the training data
cluster_2_data$cluster <- cutree(clustered, k=3)
cl2_bloom <- vector(length = 3)
#so doesn't work with subclusters 2 and 3
for(i in 1:3){
  cl2_bloom[i] <- fviz_nbclust(NbClust(cluster_2_data %>% filter(cluster == i) %>% select(ends_with(".blm"),
                              X2000nat), distance="euclidian", min.nc=2, max.nc=4, method="ward.D2", index="all"))
}


fviz_nbclust(NbClust(cluster_2_data %>% filter(cluster == 2) %>% select(ends_with(".blm"), X2000nat), distance="euclidean", 
                                min.nc=2, max.nc=4, method="ward.D2", index="all"))


hclust(cluster_2_data)
```

Now from though process, only the variables that should be included in the clustering should be the pesticides: insecticide/pesticide/thinner e.t.c

```{r}
cluster_2_data <- data_2012 %>%
  select(ends_with(".pre"),ends_with(".blm"), ends_with(".pos")) %>%
  unique()


#Clusters of whole data combined - 7
fviz_nbclust(NbClust(cluster_2_data, distance="euclidean", 
                                min.nc=2, max.nc=4, method="ward.D2", index="all"))
#pre ~ max of 4 clusters at each time point in method, but  in general if you allow
#More as this argument then more returned a better  optimal cluster number
fviz_nbclust(NbClust(cluster_2_data %>% select(ends_with(".pre")), distance="euclidean", 
                                min.nc=2, max.nc=4, method="ward.D2", index="all"))
#during
fviz_nbclust(NbClust(cluster_2_data %>% select(ends_with(".blm")), distance="euclidean", 
                                min.nc=2, max.nc=4, method="ward.D2", index="all"))
#post
fviz_nbclust(NbClust(cluster_2_data %>% select(ends_with(".pos")), distance="euclidean", 
                                min.nc=2, max.nc=4, method="ward.D2", index="all"))


#######Subsetting the data down by cluster results in the matrices being singular and therefore the code doesn't
#run as expected, this could be looked at in detail however, since there are so few data points
#it is not ideal to have a really sparse cluster set in reality

#Testing
fviz_nbclust(NbClust(cluster_2_data %>%  filter(cluster == 1) %>% select(ends_with(".pos")), distance="euclidean", 
                                min.nc=2, max.nc=4, method="ward.D2", index="all"))

cluster_2_data <- cluster_2_data %>%  mutate_all(function(x) ifelse(x == 0, runif(1, 0.00000000001,0.0000001), x))

fviz_nbclust(NbClust(cluster_2_data %>%  filter(cluster == 2) %>% select(ends_with(".pos")), distance="euclidean", 
                                min.nc=2, max.nc=4, method="ward.D2", index="all"))

bleh <- agnes(dist(cluster_2_data %>%  filter(cluster == 2) %>% select(ends_with(".blm")), method = "euclidian"), 
                   diss=TRUE, method = "ward")

cluster_2_data$cluster <- cutree(bleh, k=2)

fviz_nbclust(NbClust(cluster_2_data %>%  filter(cluster == 2) %>% select(ends_with(".blm")), distance="euclidean", 
                                min.nc=1, max.nc=3, method="ward.D2", index="all"))


a <- matrix(c(0,1,1,0), nrow=2)

#Checks
fviz_nbclust(cluster_2_data %>% filter(cluster == 2)%>% select(ends_with(".blm")),pam , method = "silhouette")


clustered <- agnes(dist(cluster_2_data, method = "euclidian"), 
                   diss=TRUE, method = "ward")
# add cluster labels to the training data
cluster_2_data$cluster <- cutree(clustered, k=3)
cl2_bloom <- vector(length = 3)
#so doesn't work with subclusters 2 and 3
for(i in 1:3){
  cl2_bloom[i] <- fviz_nbclust(NbClust(cluster_2_data %>% filter(cluster == i)%>% select(ends_with(".blm")), distance="euclidean", 
                                min.nc=1, max.nc= min(table(cluster_2_data$cluster))
                                , method="ward.D2", index="all"))
}


fviz_nbclust(NbClust(cluster_2_data , distance="euclidean", 
                                min.nc=2, max.nc=min(table(cluster_2_data$cluster))
                     , method="ward.D2", index="all"))


hclust(cluster_2_data)
```
There is just not enough data to use this clustering technique at each stage of the clustering process in a good format, so will need to ensure that sparsity is taken into consideration as a main downside to clustering methods

### Variable justification
Code to decide upon the bee variables to take into consideration for the final response variables.
```{r}
bee_correlations <- data.frame(matrix(0, nrow = 0, ncol = 0))

for(i in 1:7){
  for(j in 1:7){
  bee_correlations[i,j] <- round(cor(data_2012[5+i], data_2012[5 + j]), 2)
  }
}
a
for(i in 1:7){
 names(bee_correlations)[i] <- colnames(data_2012[5 +i])
}
bee_correlations
```

